Modeling defoliation of Pinus Radiata trees using hyperspectral remote sensing data

### 2018-08-02 01_writing/Schratz_hyperspectral.tex
methods: var imp disc: var imp

To find indices that contributed most to model performance, we used the internal variable importance measure of the \textit{xgboost} algorithm.
The score is calculated by taking the contribution of each feature for each tree in the fitted model.
The higher the score of a variable, the more important it is for the fitted model when making predictions \citep{chenXGBoostScalableTree2016}.

@@

% Performance estimates
\begin{table}[t!]
\centering
\caption[t]{Four-fold block spatial \ac{CV} performances of RR, SVM and xgboost using \ac{RMSE} as the error measure.
	Mean and standard deviation are shown.}
\begingroup\footnotesize
\begin{tabular}{llll}
	RR            & SVM           & xgboost       \\
	\hline
	59.10 (22.71) & 36.23 (15.73) & 33.26 (16.61) \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:model_comparison}
\end{table}

\begin{table}[t!]
\centering
\caption[t]{Predictive performance of \textit{xgboost} using all observations (All Observations) and observations from specific plots only (Single Plot Observations) with \ac{RMSE} as the error measure. The performance estimates for "All Observations" correspond to the fold for which the respective plot was serving as the test set. Column "Single Plot Observations", shows the mean performances at the repetition level of a spatial CV (5 folds, 5 repetitions), scored by using data of the respective plot only.}
\begingroup\footnotesize
\begin{tabular}{lll}
	\\
	Plot/Data & All Observations (Block CV) & Single Plot Observations (SpCV) \\
	\hline
	Laukiz 1  & 22.03                       & 19.18                           \\
	Laukiz 2  & 51.75                       & 17.24                           \\
	Luiando   & 13.20                       & 8.30                            \\
	Oiartzun  & 32.97                       & 14.40                           \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:supermodel_performance}
\end{table}

>>>>

To find indices that contributed most to model performance, we used the internal variable importance measure of the \textit{xgboost} algorithm.
The score is calculated by taking the contribution of each feature for each tree in the fitted model.
The higher the score of a variable, the more important it is for the fitted model when making predictions \citep{chenXGBoostScalableTree2016}.
Using the internal variable importance measure from the xgboost algorithm is quite convienient since it is automatically computed during model fit.
By contrast to other approaches such as permutation-based variable importance, it is composed out of three parts that contribute to the overall importance score:

\begin{itemize}
	\item Gain: The relative contribution of the feature to the model 
	\item Cover metric: How often a feature was selected to be the deciding feature in a tree for a specific observation 
	\item Frequency How often a feature occurs in all trees of the model 
\end{itemize}

The \textit{Gain} features is the most important one among the three. 
All measures sum up to one.
Downsides of this approach are that due to the contribution of three different parts it is complicated to understand why a specific feature was selected.
Furthermore the importance calculation approach is only valid for this algorithm and cannot be compared to others.
Nevertheless, as we only relied on the variable importance for this specific algorithm, using the internal xgboost approach was sufficient for this work.

@@

% Performance estimates
\begin{table}[t!]
\centering
\caption[t]{Four-fold block spatial \ac{CV} performances of RR, SVM and xgboost using \ac{RMSE} as the error measure.
	Mean and standard deviation are shown.}
\begingroup\footnotesize
\begin{tabular}{llll}
	RR            & SVM           & xgboost       & xgboost (7 variables) \\
	\hline
	59.10 (22.71) & 36.23 (15.73) & 33.26 (16.61) &                       \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:model_comparison}
\end{table}

\begin{table}[t!]
\centering
\caption[t]{Predictive performance of \textit{xgboost} using all observations (All Observations) and observations from specific plots only (Single Plot Observations) with \ac{RMSE} as the error measure. The performance estimates for "All Observations" correspond to the fold for which the respective plot was serving as the test set. Column "Single Plot Observations", shows the mean performances at the repetition level of a spatial CV (5 folds, 5 repetitions), scored by using data of the respective plot only.}
\begingroup\footnotesize
\begin{tabular}{llll}
	\\
	Plot/Data & All Observations/all variables (Block CV) & All Observations/7 variables (Block CV) & Single Plot Observations (SpCV) \\
	\hline
	Laukiz 1  & 22.03                                     &                                         & 19.18                           \\
	Laukiz 2  & 51.75                                     &                                         & 17.24                           \\
	Luiando   & 13.20                                     &                                         & 8.30                            \\
	Oiartzun  & 32.97                                     &                                         & 14.40                           \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:supermodel_performance}
\end{table}
